### 2016-02-19 paper.tex
Discuss the fact that MHK rotors are usually more solid

The Reference Model Project (RMP), sponsored by the US Department of Energy,
produced six marine hydrokinetic (MHK) technology point designs as reference
models (RM) to serve as non-proprietary test articles for open research
development, and to benchmark performance and costs for technology developers
\cite{Neary2014, Neary2014a}. Open-source RMP products, along with supporting
documentation, are available at the RMP website
\url{http://energy.sandia.gov/rmp} to facilitate their use in future R\&D
studies by industry, academia, and national laboratories. These products
include: Technical specifications and computer-aided design (CAD) files for each
RM device to allow exact replication for physical and numerical modeling
studies; Resource site information used to design each RM device; and references
to physical modeling data sets that can be used to validate numerical modeling
design and analysis tools.

Reference Model 2 (RM2) is a dual-rotor, vertical-axis cross-flow river turbine
that was designed to operate in a reach of the lower Mississippi River near
Baton Rouge, Louisiana \cite{Barone2011, Neary2011}. The rotor has three blades,
a relatively low solidity $Nc/(\pi D)$ or chord-to-radius ratio $c/R$, and a
preliminary analysis with Sandia's CACTUS vortex line numerical model
\cite{Murray2011} predicted a full-scale performance coefficient of 47\% at 1
m/s flow speed and a tip speed ratio $\lambda \equiv \frac{\omega R}{U_\infty} =
3.15$ \cite{Barone2011}, where $\omega$ and $R$ are the rotor's angular velocity
and radius, respectively. An initial experimental measurement campaign with a
1:15 scale RM2 rotor conducted at the Saint Anthony Falls Laboratory (SAFL) at
the University of Minnesota resulted in maximum power coefficient of
approximately 5\% at $\lambda = 2.2$. This discrepancy between numerical and
physical model performance prompted the experimental measurements presented
here, namely due to the low Reynolds number---$Re_D = U_\infty D / \nu$, where
$U_\infty$ is the free stream velocity, $D$ is the rotor diameter, and $\nu$ is
the fluid kinematic viscosity---of the 1:15 scale tests, which were performed at
$Re_D \sim 10^5$ \cite{Hill2014}.

>>>>

The Reference Model Project (RMP), sponsored by the US Department of Energy,
produced six marine hydrokinetic (MHK) technology point designs as reference
models (RMs) to serve as non-proprietary test articles for open research
development, and to benchmark performance and costs for technology developers
\cite{Neary2014, Neary2014a}. Open-source RMP products, along with supporting
documentation, are available at the RMP website
\url{http://energy.sandia.gov/rmp} to facilitate their use in future R\&D
studies by industry, academia, and national laboratories. These products
include: Technical specifications and computer-aided design (CAD) files for each
RM device to allow exact replication for physical and numerical modeling
studies; resource site information used to design each RM device; and references
to physical modeling data sets that can be used to validate numerical modeling
design and analysis tools.

Reference Model 2 (RM2) is a dual-rotor, vertical-axis cross-flow river turbine
that was designed to operate in a reach of the lower Mississippi River near
Baton Rouge, Louisiana \cite{Barone2011, Neary2011}. The rotor has three blades,
a relatively low solidity $Nc/(\pi D)$ or chord-to-radius ratio $c/R$, and a
preliminary analysis with Sandia's CACTUS vortex line numerical model
\cite{Murray2011} predicted a full-scale performance coefficient of 47\% at 1
m/s flow speed and a tip speed ratio $\lambda \equiv \frac{\omega R}{U_\infty} =
3.15$ \cite{Barone2011}, where $\omega$ and $R$ are the rotor's angular velocity
and radius, respectively. Note that the RM2 rotor solidity would be considered
moderate to high for a wind turbine, but MHK rotors are typically higher
solidity since they must withstand approximately an order of magnitude higher
torque in typical flows, necessitating relatively larger blades to meet strength
and fatigue life requirements.

An initial experimental measurement campaign with a 1:15 scale RM2 rotor
conducted at the Saint Anthony Falls Laboratory (SAFL) at the University of
Minnesota resulted in maximum power coefficient of approximately 5\% at $\lambda
= 2.2$. This discrepancy between numerical and physical model performance
prompted the experimental measurements presented here, namely due to the low
Reynolds number---$Re_D = U_\infty D / \nu$, where $U_\infty$ is the free stream
velocity, $D$ is the rotor diameter, and $\nu$ is the fluid kinematic
viscosity---of the 1:15 scale tests, which were performed at $Re_D \sim 10^5$
\cite{Hill2014}.
